### 2020-03-25 Thesis-Chapter1-Introduction.tex
Add to background

The paper serves to clarify the arguments composing debate and the nature of the technical problem. After reproducing a
brief history of encryption regulation, technology, and conflict, it introduces argument maps and threat models using
data flow diagrams. In the following chapters it analyzes exceptional access arguments in detail and creates a
multi-level threat model, broken down by the significantly different challenges posed by data at rest and data in
motion. The paper concludes with a set of policies and principles to apply to the problem in light of valid arguments
and threats.

>>>>

The paper serves to clarify the arguments composing debate and the nature of the technical problem. After reproducing a
brief U.S.-focused history of encryption regulation, technology, and conflict, it introduces argument maps and threat
models using data flow diagrams. In the following chapters it analyzes exceptional access arguments in detail and
creates a multi-level threat model, broken down by the significantly different challenges posed by data at rest and data
in motion. The paper concludes with a set of policies and principles to apply to the problem in light of valid arguments
and threats.

### 2020-03-25 Thesis-Chapter2-Background.tex
Add to background

\section{A Brief History of Encryption}
\label{sec-history}

@@

If the cryptographic primitives that compose these systems were to be compromised, the systems they compose could
collapse. If the financial system loses integrity, health information systems lose confidentiality, critical
infrastructure loses availability, government communications lose authenticity and non-repudiation, or the internet as a
whole loses integrity, inconceivably large damage would occur.

@@

In 1976, the same year Diffie and Hellmann published their research in public key cryptography, the U.S. Congress passes
the Arms Export Control Act and declares that strong cryptography is subject to export controls \cite{kehl_right_2015}.
In 1991, the U.S. Senate introduces, but does not pass, a bill mandating access to plain text contents when authorized
by law; in response, Philip Zimmermann releases PGP (Pretty Good Privacy), a public key encryption tool to secure email,
in order that strong cryptography be ``made available to the American public before it became illegal to use''
\cite{zimmermann_1996}. In 1993, the Clinton administration introduces Clipper Chip with the goal of ``providing the
public with strong cryptographic tools without sacrificing the ability of law enforcement and intelligence agencies to
access unencrypted versions of those communications'' \cite{thompson_2015}. Citing the foundational role of cryptography
in security--and the potential of human rights abuses for compromised privacy--industry and prominent technical leaders
reacted negatively to the chip \cite{kehl_right_2015} \cite{zimmermann_1996}. When security researcher Matt Blaze
discovered flaws that allowed users to subvert the Clipper Chip mechanisms \cite{blaze_protocol_1994}, the proposal
died.

@@

Two developments sewed the seeds for the Second Crypto War, and three set if off. The first seed was actually a
non-development: when the Clinton administration changed its encryption policy, the U.S. House dropped the SAFE Act
\cite{kehl_right_2015}. This meant that rather than being written in the pen of law, encryption policy was written in
the pencil of executive order. The second seed was the events of and reaction to morning of September 11, 2001.

Three developments have triggered the Second Crypto War.
(1) The 2013 Snowden revelations
(2) The mid-2010s spread of default device encryption in smartphones
(3) The late-2010s spread of default E2E encryption in messaging services

@@

@@

>>>>

\section{Historical Context}
\label{sec-history}

\subsection{Encryption History and the First Crypto War}

@@

If the cryptographic primitives that compose these systems were to be compromised, the systems they comprise could
collapse. If the financial system loses integrity, health information systems lose confidentiality, critical
infrastructure loses availability, government communications lose authenticity and non-repudiation, or the internet as a
whole loses integrity, inconceivably large damage would occur.

@@

In 1976, the same year Diffie and Hellmann published their research in public key cryptography, the U.S. Congress passed
the Arms Export Control Act and declares that strong cryptography is subject to export controls \cite{kehl_right_2015}.
In 1991, the U.S. Senate introduced, but did not pass, a bill mandating access to plain text contents when authorized by
law; in response, Philip Zimmermann released PGP (Pretty Good Privacy), a public key encryption tool to secure email, in
order that strong cryptography be ``made available to the American public before it became illegal to use''
\cite{zimmermann_1996}. In 1993, the Clinton administration introduced Clipper Chip \cite{press_1993} with the goal of
``providing the public with strong cryptographic tools without sacrificing the ability of law enforcement and
intelligence agencies to access unencrypted versions of those communications'' \cite{thompson_2015}. Citing the
foundational role of cryptography in security--and the potential of human rights abuses for compromised
privacy--industry and prominent technical leaders reacted negatively to the chip \cite{kehl_right_2015}
\cite{zimmermann_1996}. When security researcher Matt Blaze discovered flaws that allowed users to subvert the Clipper
Chip mechanisms \cite{blaze_protocol_1994}, the proposal died.

@@

\subsection{The Second Crypto War and the Current Context}

Two developments sewed the seeds for the Second Crypto War. The first was actually a non-development: when the Clinton
administration changed its encryption policy, the U.S. House dropped the SAFE Act \cite{kehl_right_2015}. This meant
that rather than being written in the pen of law, encryption policy was written in the pencil of executive order. The
second development was the events of and reaction to morning of September 11, 2001.

Three developments have triggered the Second Crypto War.
\begin{enumerate}
    \item The 2013 Snowden revelations
    \item The mid-2010s spread of default device encryption in smartphones
    \item The late-2010s spread of default E2E encryption in messaging services
\end{enumerate}

Characteristic elements of the Second Crypto War include:
\begin{enumerate}
    \item The Apple v. FBI case over access to the San Bernardino shooter's iPhone \cite{schulze_clipper_2017}
    \item A return to the Four Horsemen \cite{schneier_scaring_2019}, with a (warranted \cite{keller_internet_2019})
            focus currently on CSAM \cite{geller_2019}
    \item "Security vs Security" vs "Security vs Privacy" \cite{stalla_bourdillon_privacy_2014} \cite{schneier_2019}
    \item Pendulating emphasis on DIM/E2E and DAR/DE \cite{schneier_2019}
    \item A political landscape marked by uncertainty, extremism, misinformation, and most recently, a pandemic that
            will have unknown impacts
\end{enumerate}

That brings us to where the debate stands today. A December 2019 U.S. Senate Judiciary Committee featured the top two
U.S. parties both expressing anger over whether data should be ``beyond the reach of the law'' and expressed threats
like ``get your act together, or we will gladly get your act together for you'' \cite{geller_2019}. In March 2020, the
Republican atop the Senate Judiciary Committee introduced S.3398, the EARN IT Act of 2020, which seeks to establish a
16-chair committee with the power to revoke online platforms' liability protection for user-submitted content if the
committee determines the platform is not doing enough to fight child sexual abuse material (CSAM), a legal synonym for
child pornography \cite{graham_s3398_2020}. As illustrated by an in-depth article by the \textit{New York Times}
\cite{keller_internet_2019} that likely triggered the December 2019 hearing, the problem is enormous; however, critics
largely see the EARN IT Act as an attempt to indirectly yet effectively outlaw E2E \cite{newman_2020}
\cite{pfefferkorn_2020}.

% TODO: create timeline! https://tex.stackexchange.com/questions/196794/how-can-you-create-a-vertical-timeline

The following sections summarize the encryption regulatory environment and technical approaches to exceptional access.

@@

The following laws and regulations have been applied to cases involving cryptography and access to encrypted
information:

\begin{itemize}
    \item The Bill of Rights: Amendments I, IV, and V \cite{madison_1791}
    \item The Arms Export Control Act (AECA) of 1976 established export controls over strong encryption techniques
        introduced ITAR (International Traffic in Arms Regulations) \cite{morgan_hr13680_1976} \cite{kehl_right_2015}
    \item The Computer Fraud and Abuse Act (CFAA) of 1986 \cite{hughes_hr4718_1986} \cite{wolff_computer_2016}
    \item The Communications Assistance for Law Enforcement Act (CALEA) of 1994 \cite{edwards_hr4922_1994}
    \item \dots
\end{itemize}

The following are proposed laws and regulations that did not go into effect or have yet to be voted on:

\begin{itemize}
    \item The Clipper Chip \cite{press_1993} \cite{thompson_2015}
    \item The SAFE Act \cite{goodlatte_hr3011_1996}
    \item The EARN IT Act \cite{graham_s3398_2020}
    \item \dots
\end{itemize}

@@

The following is a very brief account of the regulatory environment in various jurisdictions around the world:

\begin{itemize}
    \item Australia: in 2018 passed its Assistance and Access bill \cite{ag_2018} \cite{newman_2018}
    \item UK: Investigatory Powers Act \cite{legislature_2016}
    \item EU: GDPR
    \item China: \cite{donahue_comparative_2018}
    \item Israel: \cite{donahue_comparative_2018}
    \item Japan, Korea? -- can't find anything
    \item \dots
\end{itemize}
